Pretty darn good control: when are approximate solutions better than approximate models

### 2023-06-30 manuscript/manuscript.Rmd
abbys comments

- $T(N_{t+1}|N_t, a_t)$: *transition operator*, a conditional distribution which describes the dynamics of the system (where $N_i\in\mathcal{S}$ are states of the environment),

- $r(N_t, a_t)$: *reward function*, the reward obtained after performing action $a_t\in\mathcal{A}$ in state $N_t$,

@@

With any fixed policy function, the agent will traverse a path 
$\tau=(N_0,\ a_0,\ N_1,\ a_1\ \dots,\ N_{t_{\text{fin.}}})$ 
sampled randomly from the distribution
\begin{align*}
  p_\pi(\tau) = 
  d(N_0) \prod_{t=0}^{ t_{\text{fin.}}-1 }
  \pi( a_t | N_t ) T( N_{t+1} | N_t, a_t ).
\end{align*}
Reinforcement learning seeks to optimize $\pi$ such that the expected rewards are maximal,
$$  
  \pi^* = \mathrm{argmax}\ \mathbb{E}_{\tau\sim p_\pi}[R(\tau)],
$$
where,
$$
  R(\tau) = \sum_{t=0}^{ t_{\text{fin.}}-1 } \gamma^t r(a_t, N_t),
$$
is the cumulative reward of path $\tau$.
The function $J(\pi):=\mathbb{E}_{\tau\sim p_\pi}[R(\tau)]$ is called the *expected return*.

@@

The reward function defining our policy optimization problem had two components.
The first was economical: the total biomass harvested over an episode.
The second seeked to reflect conservation goals: if a near-extinction event occurred at time $t$, the episode was ended early and a negative reward of $-100/t$ was awarded.
This reward function balanced the extractive motivation of the fishery with conservation goals which went beyond the scope of long-term sustainable harvests.

@@

```{r}
#| label: ephistograms
#| fig.cap: "Hisotgrams of episode lengths and rewards for the four different management strategies considered. Only the first 50 evaluation episodes (from a total of 100) were included, for ease of visualization."
#| out.width: "6in"
esc_len <- esc_ends %>%
  filter(rep<50) %>%
  ggplot(
    mapping = aes(x = rep, y = t)
  ) +
  geom_col() +
  labs(x="", y="ep. duration")
msy_len <- msy_ends %>%
  filter(rep<50) %>%
  ggplot(
    mapping = aes(x = rep, y = t)
  ) +
  geom_col() +
  labs(x="", y="")
ppo_len <- ppo_ends %>%
  filter(rep<50) %>%
  ggplot(mapping = aes(x = rep, y = t)) +
  geom_col() +
  labs(x="", y="")
gpp_len <- gpp_ends %>%
  filter(rep<50) %>%
  ggplot(
    mapping = aes(x = rep, y = t)
  ) +
  geom_col() +
  labs(x="", y="")

@@

Here we summarize the procedure used to interpolate the PPO policy (visualized in Fig. \ref{fig:ppo-policy}).
We use the *GaussianProcessRegressor* object of the *sklearn* Python library with a kernel given by
\begin{align*}
  \text{RBF}(\text{length scale = 10}) + \text{WhiteNosie}(\text{noise level = 0.1}).
\end{align*}
This interpolation method is applied to scatter data of the PPO policy evaluated on 3 different grids on $(X,Y,Z)$ states:
$G_X$, a $51\times 5 \times 5$ grid; $G_Y$, a $5\times 51 \times 5$ grid; and $G_Z$, a $5\times 5 \times 51$ grid.
This combination of grids was used instead of a single dense grid in order to reduce the computational intensity of the interpolation procedure.
For $G_X$, the 5 values for $Y$ and $Z$ were varied in a ‘‘popular window,’’ i.e. episode time-series data was used to determine windows of $Y$ and $Z$ values which were most likely. 
The grids $G_Y$ and $G_Z$ were generated in a similar fashion, *mutatis mutandis*.\footnote{
The raw dataset is found at the data/results\_data/2FISHERY/RXDRIFT sub-directory in the repository with the source code and data linked above.
Scatter plots visualizing this policy are shown in App. B.
}
The length scale and noise level values of this kernel were chosen arbitrarily---no hyperparameter tuning was needed to produce satisfactory interpolation, as will be shown in the results section.

>>>>

- $T(N_{t+1}|N_t, a_t, t)$: *transition operator*, a conditional distribution which describes the dynamics of the system (where $N_i\in\mathcal{S}$ are states of the environment),\footnote{
Transition operators are commonly discussed without having a direct time-dependence for simplicity, but the inclusion of $t$ as an argument to $T$ does not alter the structure of the learning problem appreciably.
}

- $r(N_t, a_t, t)$: *reward function*, the reward obtained after performing action $a_t\in\mathcal{A}$ in state $N_t$,

@@

With any fixed policy function, the agent will traverse a path 
$\tau=(N_0,\ a_0,\ N_1,\ a_1\ \dots,\ N_{t_{\text{fin.}}})$ 
sampled randomly from the distribution
\begin{align*}
  p_\pi(\tau) = 
  d(N_0) \prod_{t=0}^{ t_{\text{fin.}}-1 }
  \pi( a_t | N_t ) T( N_{t+1} | N_t, a_t, t ).
\end{align*}
Reinforcement learning seeks to optimize $\pi$ such that the expected rewards are maximal,
$$  
  \pi^* = \mathrm{argmax}\ \mathbb{E}_{\tau\sim p_\pi}[R(\tau)],
$$
where,
$$
  R(\tau) = \sum_{t=0}^{ t_{\text{fin.}}-1 } \gamma^t r(a_t, N_t, t),
$$
is the cumulative reward of path $\tau$.
The function $J(\pi):=\mathbb{E}_{\tau\sim p_\pi}[R(\tau)]$ is called the *expected return*.

@@

The reward function defining our policy optimization problem had two components.
The first was economical: the total biomass harvested over an episode.
The second seeked to reflect conservation goals: if a near-extinction event occurred at time $t$, the episode was ended early and a negative reward of $-100/t$ was awarded.
This reward function balanced the extractive motivation of the fishery with conservation goals which went beyond the scope of long-term sustainable harvests commonly used in fishery management.

@@

```{r}
#| label: ephistograms
#| fig.cap: "Hisotgrams of episode lengths and rewards for the four different management strategies considered. Only the first 50 evaluation episodes (from a total of 100) were included, for ease of visualization. From left to right, the four management strategies compared are CEsc, CMort, PPO, and PPO+GP."
#| out.width: "6in"
esc_len <- esc_ends %>%
  filter(rep<50) %>%
  ggplot(
    mapping = aes(x = rep, y = t)
  ) +
  geom_col() +
  labs(x="", y="ep. duration") +
  theme(plot.title = element_text(hjust = 0.5))+
  ggtitle("CEsc")
msy_len <- msy_ends %>%
  filter(rep<50) %>%
  ggplot(
    mapping = aes(x = rep, y = t)
  ) +
  geom_col() +
  labs(x="", y="") +
  theme(plot.title = element_text(hjust = 0.5))+
  ggtitle("CMort")
ppo_len <- ppo_ends %>%
  filter(rep<50) %>%
  ggplot(mapping = aes(x = rep, y = t)) +
  geom_col() +
  labs(x="", y="")+
  theme(plot.title = element_text(hjust = 0.5))+
  ggtitle("PPO")
gpp_len <- gpp_ends %>%
  filter(rep<50) %>%
  ggplot(
    mapping = aes(x = rep, y = t)
  ) +
  geom_col() +
  labs(x="", y="")+
  theme(plot.title = element_text(hjust = 0.5))+
  ggtitle("PPO+GP")

@@

Here we summarize the procedure used to interpolate the PPO policy (visualized in Fig. \ref{fig:ppo-policy}).
We use the *GaussianProcessRegressor* object of the *sklearn* Python library with a kernel given by
\begin{align*}
  \text{RBF}(\text{length scale = 10}) + \text{WhiteNoise}(\text{noise level = 0.1}).
\end{align*}
This interpolation method is applied to scatter data of the PPO policy evaluated on 3 different grids on $(X,Y,Z)$ states:
$G_X$, a $51\times 5 \times 5$ grid; $G_Y$, a $5\times 51 \times 5$ grid; and $G_Z$, a $5\times 5 \times 51$ grid.
This combination of grids was used instead of a single dense grid in order to reduce the computational intensity of the interpolation procedure.
For $G_X$, the 5 values for $Y$ and $Z$ were varied in a ‘‘popular window,’’ i.e. episode time-series data was used to determine windows of $Y$ and $Z$ values which were most likely. 
The grids $G_Y$ and $G_Z$ were generated in a similar fashion, *mutatis mutandis*.\footnote{
The raw dataset is found at the data/results\_data/2FISHERY/RXDRIFT sub-directory in the repository with the source code and data linked above.
Scatter plots visualizing this policy are shown in App. B.
}
The length scale and noise level values of this kernel were chosen arbitrarily---no hyperparameter tuning was needed to produce satisfactory interpolation, as will be shown in the results section.
